### 2016-06-23 filters.tex
responding to JK's comments

\heading{FP-security of hash-based filters.} \tsnote{NY give it in
the private hash setting (sort of); but, intuitively, it should be
hard to find false-positives even if you know that MD5 or SHA-x are
the hash functions.  Not totally sure how to capture this intuition
nicely.} \jnote{Not sure I agree. First of all, if $\epsilon$ is not
negligible then a random element is a false positive with noticeable
probability. Beyond that, I think that intuitively it makes sense
that if you want false positives to be hard to find then you need to
make the hash functions secret.} The traditional notion of soundness
for a Bloom filter captures the difficulty of finding false
positives when the attacker is non-adaptive. Following NY, we
consider soundness against adaptive adversaries. But we further
generalize to allow for settings in which the attacker may have some
a priori knowledge about the contents of~$S$. (The Niedermaier et
al.~\cite{xxx} attacks, for example, exploit knowledge about the
distribution of German last names.)   In particular, our notions
take as input an explicit distribution $\distr{U}{n}$ over~$[U]^n$.

\jknote{We also want to capture the fact that finding the first
false positive should not make it any easier to find subsequent
false positives. This is not really captured by the Naor paper,
either (at least not directly, though it may end up being implied by
their definition).}

@@

\jnote{How do we want to define a ``secure'' Bloom filter (even
informally)? Do we want to say that the probability of finding a
false positive is negligible? Or that if we set the FP rate to
$\epsilon$ then the best an attacker can do with $q$ hash queries is
$1-(1-\epsilon)^q$ plus negligible? One thing that is nice about the
NY notion is that you get a ``clean'' definition of security: no
matter how many FPs the attacker has found so far, the probability
that it can output a fresh FP is still $\epsilon$.}

>>>>

\heading{FP-security of hash-based filters.} \tsnote{NY give it in
the private hash setting (sort of); but, intuitively, it should be
hard to find false-positives even if you know that MD5 or SHA-x are
the hash functions.  Not totally sure how to capture this intuition
nicely.} \jnote{Not sure I agree. First of all, if $\epsilon$ is not
negligible then a random element is a false positive with noticeable
probability. Beyond that, I think that intuitively it makes sense
that if you want false positives to be hard to find then you need to
make the hash functions secret.} \tsnote{I don't think it makes sense
to talk about asymptotics or negligibility, for exactly the reason you
state. I think the key point here is: what is ``hard''?   The designer decides what FP rate is
acceptable for a given application, a given set of efficiency
constraints, the side-effects of a FP, etc.  For a fixed FP target,
she optimizes the filter size and/or the number of hash functions for
her setting. Whether the target FP rate is 0.01 or $2^{-80}$, we should be able to say (in the ROM, at least)
that it will take this-many hash computations, and this many on-line
set-membership queries, to find the first FP.  In the standard model,
with public hash functions, it is clearly more difficult to make
formal claims.  Even so, in practice I expect it would be quite expensive to
find a string~$x$ such that $(h_1(x),h_2(x),\ldots,h_k(x))$
falls in a given set of tuples when $h_i(x)=\mathrm{truncate}(\mathrm{SHA256}(\langle i \rangle
\concat x), m)$, say. (Even more so if my setting admits relatively
expensive hash computations, e.g. SCRYPT.)  BTW, your suggestion to
have a FP notion with secret~$\mathrm{S}$ is interesting here. }

The traditional notion of soundness
for a Bloom filter captures the difficulty of finding false
positives when the attacker is non-adaptive. Following NY, we
consider soundness against adaptive adversaries. But we further
generalize to allow for settings in which the attacker may have some
a priori knowledge about the contents of~$S$. (The Niedermaier et
al.~\cite{xxx} attacks, for example, exploit knowledge about the
distribution of German last names.)   In particular, our notions
take as input an explicit distribution $\distr{U}{n}$ over~$[U]^n$.

\jknote{We also want to capture the fact that finding the first
false positive should not make it any easier to find subsequent
false positives. This is not really captured by the Naor paper,
either (at least not directly, though it may end up being implied by
their definition).} \tsnote{See my comment in the intro.}

@@

\jnote{How do we want to define a ``secure'' Bloom filter (even
informally)? Do we want to say that the probability of finding a
false positive is negligible? Or that if we set the FP rate to
$\epsilon$ then the best an attacker can do with $q$ hash queries is
$1-(1-\epsilon)^q$ plus negligible?} 
\tsnote{Formally, I don't think we can (and still be rooted in
  reality.)  Even informally it's hard, since the fact is that the
  designer decides what is ``secure'' for her setting.  I think we
  just say something like ``A bloom filter is
  $(t,q_h,q_m,\epsilon)$-secure if no attacker running in time at
  most~$t$, asking at most~$q_h$ hash queries and $q_m$ set-membership
  queries, has probability more than~$\epsilon$ of finding a FP.''  (If
you like, we can throw a paramter~$n$ in there, for the number of FPs
it must find.)  Perhaps informally we say that a bloom filter is
secure if, for ``reasonable'' $t,q_h,q_m$, the value of~$\epsilon$ is
not ``significantly larger'' than the designer's target.}

\jnote{One thing that is nice about the
NY notion is that you get a ``clean'' definition of security: no
matter how many FPs the attacker has found so far, the probability
that it can output a fresh FP is still $\epsilon$.}

### 2016-06-23 intro.tex
responding to JK's comments

\section{Introduction}
We revisit the formalization of Bloom filters~\cite{xxx} and closely
related data structures as primitives with explicit security goals.
Recently, Naor and Yogev~\cite{xxx} (hereafter NY) gave such a
formalization, providing syntax for a generalized abstraction of
Bloom filters, as well as casting the classical Bloom filter
soundness condition as a security experiment admiting adaptive
adversaries.  To briefly summarize, a Bloom filter is a data
structure for compactly representing a set~$\mathcal{S}$ in a way
that supports constant-time responses to set-membership queries. The
representation of~$\mathcal{S}$ is an bit-array~$M$ of length~$m$,
all initially set to~0.  To insert $x \in \mathcal{S}$ into~$M$, one
computes indices $h_1(x),h_2(x),\ldots,h_k(x)$ where each $h_i
\colon \mathcal{U} \to \{0,1,\ldots m-1\}$ maps from a
universe~$\mathcal{U}$ (which contains $\mathcal{S}$) to indices
of~$M$.  The bits at positions $h_1(x),h_2(x),\ldots,h_k(x)$ are set
to 1; if a bit is already 1, due to inserting some previous element
of~$\mathcal{S}$, then it is left alone.  After all elements
of~$\mathcal{S}$ have been inserted, the resulting bit-array~$M$ is
taken as the representation.  To test if $y \in \mathcal{S}$, one
checks if \[(M[h_1(y)]=1 \wedge M[h_2(y)]=1 \wedge \cdots \wedge
M[h_k(y)]=1),\] returning ``yes'' if so, and ``no'' otherwise. Bloom
filters are correct, in the sense that if $y \in \mathcal{S}$, the
set-membership test always returns ``yes''.  The classical soundness
condition asserts that for all $\mathcal{S}$ and any $y
\notin\mathcal{S}$, we have $\Prob{h_1,h_2,\ldots,h_k \getsr
\mathrm{Func}(\mathcal{U},[m])\colon M[h_1(y)]=1 \wedge \cdots
\wedge M[h_k(y)]=1} \leq \epsilon$ where $\epsilon$, the
false-positive rate, is a parameter of the Bloom filter.  (The
values of~$\epsilon,m,k$, and $|\mathcal{S}|$ are related in a
well-known way.) NY recast this soundness condition as a game in
which an adversary~$A$ is given~$\mathcal{S}$ and an oracle for
set-membership queries, and it attempts to find a false positive.
The fact that~$A$ can \emph{adapt} its queries and its output based
on what it learns from the set-membership oracle is the key
definitional distribution of~NY.

@@

Syntactically, NY says a Bloom filter (with a steady representation)
is a pair of algorithms $\mathbf{B}=(\mathbf{B}_1, \mathbf{B}_2)$.
The setup algorithm~$\mathbf{B}_1$ is randomized, and on input a
set~$\mathcal{S}$ it returns a representation~$M$ (from some
undefined, implicit set of representations).  The query
algorithm~$\mathcal{B}_2$ is deterministic, and on input a
representation~$M$ and an element~$x \in \mathcal{U}$, it returns a
bit.  For correctness they make the classical demand, that for all
$x \in \mathcal{S}$ we have
$\Prob{\mathbf{B}_2(\mathbf{B}_1(\mathcal{S}),x)=1}=1$.  For
soundness, adversary~$A$ plays the following game,
``Adversarial-resilient Bloom Filter with a steady representation''.
First, $M \getsr \mathbf{B}_1(\mathcal{S})$. \jnote{Where does the
set come from? Is it random? Adversarially chosen?} Then the
adversary is given the set~$\mathcal{S}$ and may make queries
$x_1,x_2,\ldots,x_q$ to an oracle for $\mathbf{B}_2(M,\cdot)$.  At
the end of its execution, the adversary outputs~$x^*$ and wins the
game if (1) $x^* \not\in \mathcal{S} \cup \{x_1,x_2,\ldots,x_q\}$,
and (2) $\mathbf{B}_2(M,x^*)=1$.

First, we note that the security notion forbids~$A$ to win if it
outputs $x^* \in \{x_1,x_2,\ldots,x_q\}$, even if one of these was a
false positive. \jnote{We discussed this on the phone, but let me
put it in writing here. I think this is actually a reasonable choice
because it captures the notion that ``even after finding some false
positives, it remains hard to find additional false positives.''
Note that in practice finding false positives is ``easy'' since
$\epsilon$ might be $1/1000$ rather than negligible.}
%
Second, we note that the NY syntax does not capture the size of~$M$
as a parameter \jnote{I think this is also reasonable; typically
security is defined without regard for efficiency}; the surrounding
text does say that $\mathbf{B}_1$ ``outputs a compressed
representation [of $\mathcal{S}$] $\mathbf{B}_1(\mathcal{S})=M$''
but what exactly this means is never addressed.  The formal syntax
admits $\mathbf{B}_1(\mathcal{S})=\mathcal{S}$ and
$\mathbf{B}_2(\mathcal{S},x)=1 \Leftrightarrow x \in \mathcal{S}$,
which is also trivially secure by their definition.  These are both
small, but important complaints.  But a larger complaint, and one
that is much harder to see, is that that it is not clear that the NY
formalism actually captures real world Bloom filters.

@@

\jnote{FWIW, I think the classical description of Bloom filters has
the same problem. I.e., if the hash functions are not fixed, then
their description should be counted as part of the length of the
data structure. On the other hand, if they are fixed, then the
false-positive rate is not well-defined (since there is no
probability space to speak of). So in that sense, NY just carried
over the problem from the classical literature.}

On the other hand, assume the hash functions are implicitly
hard-coded into~$\mathbf{B}_1$ (and $\mathbf{B}_2$). Then we can
treat~$\mathbf{B}_1$ as deterministic, since it will simply carry
out the steps required to populate an array using these hash
functions, and output the array~$M$ as the representation.  If the
descriptions of~$\mathbf{B}_1,\mathbf{B}_2$ are known to the
adversary, who is also provided with~$\mathcal{S}$, then the
adversary can compute~$M$ itself, as well as test for false
positives itself. So it isn't clear how the security notion helps to
reason about difficulty of finding a false-positive. Moreover, NY
stresses (at least once) that the representation is \emph{not} given
to the adversary in their notions.  So it seems unlikely that they
intended for $h_1,h_2,\ldots,h_k$ to be hard-coded into the
algorithms. \jnote{Perhaps it is reasonable to consider a definition
in which the set is not given to the adversary?}

@@

We generalize this syntax to hash-based filters that are mutable
\jnote{I find ``mutable'' a slightly awkward term,  maybe because it
makes me think of mutability in the context of programming
languages. In much data structures literature, this is called the
``dynamic'' case, in contrast to the ``static'' case.}, in the sense
that they allow for updating of the representation. This allows us
to capture counting Bloom filters, spectral Bloom filters, stable
Bloom filters, count-min sketches, and other compact set
representation data structures...

>>>>

\section{Introduction}
We revisit the formalization of Bloom filters~\cite{xxx} and closely
related data structures as primitives with explicit security goals.
Recently, Naor and Yogev~\cite{xxx} (hereafter NY) gave such a
formalization, providing syntax for a generalized abstraction of
Bloom filters, as well as casting the classical Bloom filter
soundness condition as a security experiment admiting adaptive
adversaries.  To briefly summarize, a Bloom filter is a data
structure for compactly representing a set~$\mathcal{S}$ in a way
that supports constant-time responses to set-membership queries. The
representation of~$\mathcal{S}$ is an bit-array~$M$ of length~$m$,
all initially set to~0.  To insert $x \in \mathcal{S}$ into~$M$, one
computes indices $h_1(x),h_2(x),\ldots,h_k(x)$ where each $h_i
\colon \mathcal{U} \to \{0,1,\ldots m-1\}$ maps from a
universe~$\mathcal{U}$ (which contains $\mathcal{S}$) to indices
of~$M$.  The bits at positions $h_1(x),h_2(x),\ldots,h_k(x)$ are set
to 1; if a bit is already 1, due to inserting some previous element
of~$\mathcal{S}$, then it is left alone.  After all elements
of~$\mathcal{S}$ have been inserted, the resulting bit-array~$M$ is
taken as the representation.  To test if $y \in \mathcal{S}$, one
checks if \[(M[h_1(y)]=1 \wedge M[h_2(y)]=1 \wedge \cdots \wedge
M[h_k(y)]=1),\] returning ``yes'' if so, and ``no'' otherwise. Bloom
filters are correct, in the sense that if $y \in \mathcal{S}$, the
set-membership test always returns ``yes''.  The classical soundness
condition asserts that for all $\mathcal{S}$ and any $y
\notin\mathcal{S}$, we have $\Prob{h_1,h_2,\ldots,h_k \getsr
\mathrm{Func}(\mathcal{U},[m])\colon M[h_1(y)]=1 \wedge \cdots
\wedge M[h_k(y)]=1} \leq \epsilon$ where $\epsilon$, the
false-positive rate, is a parameter of the Bloom filter.  (The
values of~$\epsilon,m,k$, and $|\mathcal{S}|$ are related in a
well-known way.) NY recast this soundness condition as a game in
which an adversary~$A$ is given~$\mathcal{S}$ and an oracle for
set-membership queries, and it attempts to find a false positive.
The fact that~$A$ can \emph{adapt} its queries and its output based
on what it learns from the set-membership oracle (but, as we will see,
not upon the hash functions) is the key definitional contribution of~NY.

@@

Syntactically, NY says a Bloom filter (with a steady representation)
is a pair of algorithms $\mathbf{B}=(\mathbf{B}_1, \mathbf{B}_2)$.
The setup algorithm~$\mathbf{B}_1$ is randomized, and on input a
set~$\mathcal{S}$ it returns a representation~$M$ (from some
undefined, implicit set of representations).  The query
algorithm~$\mathcal{B}_2$ is deterministic, and on input a
representation~$M$ and an element~$x \in \mathcal{U}$, it returns a
bit.  For correctness they make the classical demand, that for all
$x \in \mathcal{S}$ we have
$\Prob{\mathbf{B}_2(\mathbf{B}_1(\mathcal{S}),x)=1}=1$.  For
soundness, adversary~$A$ plays the following game,
``Adversarial-resilient Bloom Filter with a steady representation''.
First, $M \getsr \mathbf{B}_1(\mathcal{S})$. \jnote{Where does the
set come from? Is it random? Adversarially chosen?}
\tsnote{In their formalism, it is implicitly an input to the security experiment.
Security is universally quantified over sets~$\mathcal{S}$.  This
would seem to be equivalent to letting the adversary
pick~$\mathcal{S}$ (prior to the hash functions being picked), with
the time to compute and write out~$\mathcal{S}$ folded into the
adversary's running time. } 
Then the
adversary is given the set~$\mathcal{S}$ and may make queries
$x_1,x_2,\ldots,x_q$ to an oracle for $\mathbf{B}_2(M,\cdot)$.  At
the end of its execution, the adversary outputs~$x^*$ and wins the
game if (1) $x^* \not\in \mathcal{S} \cup \{x_1,x_2,\ldots,x_q\}$,
and (2) $\mathbf{B}_2(M,x^*)=1$.

First, we note that the security notion forbids~$A$ to win if it
outputs $x^* \in \{x_1,x_2,\ldots,x_q\}$, even if one of these was a
false positive. 
\jnote{We discussed this on the phone, but let me
put it in writing here. I think this is actually a reasonable choice
because it captures the notion that ``even after finding some false
positives, it remains hard to find additional false positives.''
Note that in practice finding false positives is ``easy'' since
$\epsilon$ might be $1/1000$ rather than negligible.} 
\tsnote{For me the question is: is security violated once the first FP is found, or
the second?  My answer, which is in line with the traditional
soundness notion, is that the first FP is a security violation
even if finding it isn't ``hard'' by cryptographic standards.
But I do appreciate your point.  A natural goal is to say that, in some sense,
\emph{adaptivity doesn't make it any easier} to find a FP.  Adapting
to what, though?  In the traditional notion, $x$~is picked before the
coins of~$\mathbf{B}_1$ are flipped, i.e., before the hash functions
are sampled.  The direct non-adaptive generalization is to say that
$x_1,x_2,\ldots,x_q$ are all picked up front.  The direct
\emph{adaptive} generalization would allow $x_1,x_2,\ldots,x_q$ to
depend in the coins of~$\mathbf{B}_1$, i.e., the hash functions.  The
NY definition captures only indirect dependence: adversaries that can adapt to set-membership queries,
but \emph{not} (directly) to the input-output behaviors of the hash
functions. }
\tsnote{Perhaps a more obvious way to capture the idea that ``finding
  some FPs doesn't help find additional ones'' is to parameterize my notion(s) with a number~$n$, and say
that that the adversary wins iff it outputs~$n$ FPs (these may have
been previously queried to the set-membership oracle).  
Note that in the ROM (and probably in practice) finding~$n>1$ FPs will
require~$n$ times the work; finding the first won't help in finding
the second.} 
%\tsnote{Here's an off-the-wall idea.  What if the basic FP security
%notion gave the adversary an oracle for set-membership (and the hash functions, depending on the setting), and after interacting with these
%it is challenged with either $x_0$ or $x_1$, where $x_0$ is a random
%point from the set of non-FPs, and $x_1$ is a random point from the
%set of FPs?  (No oracles after being challenged.)  The adversary has
%to decide which it received.}

Second, we note that the NY syntax does not capture the size of~$M$
as a parameter; 
\jnote{I think this is also reasonable; typically
security is defined without regard for efficiency}
\tsnote{My complaint isn't about efficiency. It's about the fact that,
by not including $|M|$ in the definition, the trivial scheme I give
(just below) is allowed, and is secure.  Anyway, as I say at the end
of this paragraph, this is a small(er) complaint that what follows}. The surrounding
text does say that $\mathbf{B}_1$ ``outputs a compressed
representation [of $\mathcal{S}$] $\mathbf{B}_1(\mathcal{S})=M$''
but what exactly this means is never addressed.  The formal syntax
admits $\mathbf{B}_1(\mathcal{S})=\mathcal{S}$ and
$\mathbf{B}_2(\mathcal{S},x)=1 \Leftrightarrow x \in \mathcal{S}$,
which is also trivially secure by their definition.  These are both
small, but important complaints.  But a larger complaint, and one
that is much harder to see, is that that it is not clear that the NY
formalism actually captures real world Bloom filters.

@@

\jnote{FWIW, I think the classical description of Bloom filters has
the same problem. I.e., if the hash functions are not fixed, then
their description should be counted as part of the length of the
data structure. On the other hand, if they are fixed, then the
false-positive rate is not well-defined (since there is no
probability space to speak of). So in that sense, NY just carried
over the problem from the classical literature.}
\tsnote{One can suggest that it \emph{should} be counted in the classical
  description, but the fact is that it isn't :)  And in the 30+ years
  since the original paper, no one saw fit to correct this?  From an
  algorithmic perspective, I think the classical approach makes
  sense.  The questions being answered are: how many bits do I need to
  store a representation of the set for a given FP rate, and how does this compare to the
  combinatorially computed lower-bound?  The description length of the
hash functions may be bigger or smaller, but the number of bits you
need to hold the set representation is independent of that.  (Imagine
that you have long random bitstrings in the set~$\mathcal{S}$.  Then
whether your hash functions simply truncate, or they are random
oracles, the size of the representation as a function of the target FP
rate doesn't change.  Yet the description lengths of the hash
functions are drastically different.)}

On the other hand, assume the hash functions are implicitly
hard-coded into~$\mathbf{B}_1$ (and $\mathbf{B}_2$). Then we can
treat~$\mathbf{B}_1$ as deterministic, since it will simply carry
out the steps required to populate an array using these hash
functions, and output the array~$M$ as the representation.  If the
descriptions of~$\mathbf{B}_1,\mathbf{B}_2$ are known to the
adversary, who is also provided with~$\mathcal{S}$, then the
adversary can compute~$M$ itself, as well as test for false
positives itself. So it isn't clear how the security notion helps to
reason about difficulty of finding a false-positive. Moreover, NY
stresses (at least once) that the representation is \emph{not} given
to the adversary in their notions.  So it seems unlikely that they
intended for $h_1,h_2,\ldots,h_k$ to be hard-coded into the
algorithms. 
\jnote{Perhaps it is reasonable to consider a definition
in which the set is not given to the adversary?}
\tsnote{A FP security notion?  Yes, perhaps so.  It's probably a more
  realistic notion anyway -- is a spammer really going to know my
  whitelist a priori?  (If he does, I probably have bigger problems in
my system than receiving new spam email.)  Or set of URLs cached by a
CDN server?}

@@

We generalize this syntax to hash-based filters that are mutable
\jnote{I find ``mutable'' a slightly awkward term,  maybe because it
makes me think of mutability in the context of programming
languages. In much data structures literature, this is called the
``dynamic'' case, in contrast to the ``static'' case.} \tsnote{I
adopted this term after reading it in some bloom filter
paper(s). Ironically, I liked it for the same reason you think it is
awkward :)  My PL association to the term is that an immutable object
is one whose state cannot be modified after it is created, and a
mutable object is the opposite.  ButI don't really care too much.}, in the sense
that they allow for updating of the representation. This allows us
to capture counting Bloom filters, spectral Bloom filters, stable
Bloom filters, count-min sketches, and other compact set
representation data structures...
